---
title: "Activity"
output: github_document
Author: Raoul Noumbissi
Date : April 1, 2020
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Tracking the activity of an individual for each day for the months of October and November at 5 minute intervals.
===================================================================================================================

Load Activity.csv here:
```{r dataset}
Activity<- read.csv("Activity.csv")
Activity$Date<-Activity$date
Activity<-Activity[,c(4,3,1)]
```

Here is a histogram followed by a summary of the total steps taken per day

```{r histogram1 steps per Day } 

steps_per_Day<-tapply(Activity$steps,Activity$Date,sum)
hist(steps_per_Day)

xt<-summary(steps_per_Day)
print(xt, type="html")

mean<-summary(steps_per_Day)["Mean"]
median<-summary(steps_per_Day)["Median"]

```
We can report that the test subject took on average `r mean` steps per day. The median value for the number of steps per day is `r median`.

Create a time series plot of the average number of steps taken.
First, remove all the raws with NA values, then find the average steps taken for each interval 0 through 235.
The maximum average step values is at interval 835.
```{r remove NA } 
Activity_without_NA<-subset(Activity,!is.na(Activity$steps))
Avg_steps_by_interval<-tapply(Activity_without_NA$steps,(Activity_without_NA$interval),mean)
with(Activity_without_NA,plot(unique(interval),Avg_steps_by_interval,type="l"))
abline(v=835,col="red")

```

Here, replace each NA steps value for each time interval by its corresponding average step value for that interval.
First, get a data frame of all steps  value = NA
```{r retrieve NA } 
Activity_NA<-subset(Activity,is.na(Activity$steps))
number_Na_values<-nrow(Activity_NA)
```
There are `number_Na_values` NA values in the Activity data frame

Second, replace steps column in the **Activity_NA** data frame by 8 times the Avg_steps_by_interval vector since there are a days in the data frame.
Raw bind the two data frames  **Activity_NA** and **Activity_without_NA**, then use the *arrange* function of the **lubridate** library to order the combined data frame in ascending order of date.
```{r merge data} 
library(dplyr)
Activity_NA$steps<-rep(Avg_steps_by_interval,8)

Activity_new<-rbind(Activity_without_NA,Activity_NA)
Activity_new<-arrange(Activity_new,Activity_new$Date)
```

Here is a histogram followed by a summary of the total steps taken per day with the new Activity dataf frame where all missing values have been imputed.
```{r histogram2 steps per Day } 
new_steps_per_Day<-tapply(Activity_new$steps,Activity_new$Date,sum)
hist(new_steps_per_Day)

xt<-summary(new_steps_per_Day)
print(xt, type="html")

mean_val<-summary(new_steps_per_Day)["Mean"]
median_val<-summary(new_steps_per_Day)["Median"]

```
The mean and median steps per day are `r mean_val` and `r median_val`. Those values are similar to the mean and median values obtained earlier.

plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r line plot } 
Activity_new["Weekday"]<-weekdays(as.Date(as.character(Activity_new$Date)))
Activity_new$Weekday<-as.factor(Activity_new$Weekday)

Activity_new_weekend<-subset(Activity_new,Weekday %in% c("Saturday","Sunday"))
Activity_new_weekday<-subset(Activity_new,Weekday %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))

Avg_steps_by_interval_weekday<-tapply(Activity_new_weekday$steps,(Activity_new_weekday$interval),mean)
Avg_steps_by_interval_weekend<-tapply(Activity_new_weekend$steps,(Activity_new_weekend$interval),mean)


par(mfrow=c(2,1), mar=c(4,4,2,1))
with(Activity_new_weekend,plot(unique(interval),Avg_steps_by_interval_weekend,type="l",main="weekend", xlab ="" ,ylab="Number 0f steps"))
with(Activity_new_weekday,plot(unique(interval),Avg_steps_by_interval_weekday,type="l",main="weekday", xlab ="Interval" ,ylab="Number of steps")) 

```

